---
title: "lab3a_lauren.Rmd"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Starting with the Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
options(max.print=1000)
```

```{r}
download.file(url = "https://ndownloader.figshare.com/files/2292169",
              destfile = "data_raw/portal_data_joined.csv")
```

```{r}
surveys <- read.csv("data_raw/portal_data_joined.csv")
```

```{r}
surveys
```

```{r}
head(surveys)
```

```{r}
tail(surveys)
```

```{r}
View(surveys)
```

```{r}
str(surveys)
```

```{r}
str(as.factor(surveys$species_id))
```

## Inspecting data.frame objects
## Size
```{r}
dim(surveys)
nrow(surveys)
ncol(surveys)
```

## Content
```{r}
head(surveys)
tail(surveys)
```

## Names
```{r}
names(surveys)
colnames(surveys)
rownames(surveys)
```

## Summary
```{r}
str(surveys)
summary(surveys)
```

## Challenge 1: Based on the output of str(surveys), can you answer the following questions?

## What is the class of the object surveys?
## -Answer: data.frame
## How many rows and how many columns are in this object?
## - Answer: 34786 rows and 12 collumns
## How many species have been recorded during these surveys?
## Answer: 48 Note: You permitted me to skip this part because I cannot produce the normal results. The results I got from the summary(survey) command was the one I would have gotten if I used Tidyverse. 



## Indexing and Subsetting Data Frames

# first element in the first column of the data frame (as a vector)
```{r}
surveys[1, 1]
```

# first element in the 6th column (as a vector)
```{r}
surveys[1, 6]
```

# first column of the data frame (as a vector)
```{r}
surveys[, 1] 
```

# first column of the data frame (as a data.frame)
```{r}
surveys[1]
```

# first three elements in the 7th column (as a vector)
```{r}
surveys[1:3, 7]
```

# the 3rd row of the data frame (as a data.frame)
```{r}
surveys[3, ]  
```

# equivalent to head_surveys <- head(surveys)
```{r}
head_surveys <- surveys[1:6, ] 
```

```{r}
1:10
10:1
surveys[, -1]          # The whole data frame, except the first column
surveys[-(7:34786), ] # Equivalent to head(surveys)
```

```{r}
surveys["species_id"]       # Result is a data.frame
surveys[, "species_id"]     # Result is a vector
surveys[["species_id"]]     # Result is a vector
surveys$species_id          # Result is a vector
```



## Challenge 2

## 1. Create a data.frame (surveys_200) containing only the data in row 200 of the surveys dataset.

```{r}
surveys_200 <- surveys[200,]
```

## 2. Notice how nrow() gave you the number of rows in a data.frame? Use that number (34786) to pull out just that last row in the data frame.Compare that with what you see as the last row using tail() to make sure it’s meeting expectations.Pull out that last row using nrow() instead of the row number.

```{r}
nrow(surveys)

surveys[34786,]

tail(surveys)

nrow(surveys_last)

```

## 3. Create a new data frame (surveys_last) from that last row.
```{r}
surveys_last <- surveys[34786,]
```

## 4. Use nrow() to extract the row that is in the middle of the data frame. Store the content of this row in an object named surveys_middle.
```{r}

surveys[17393,]

surveys_middle <- surveys[17393,]
```


## 5. Combine nrow() with the - notation above to reproduce the behavior of head(surveys), keeping just the first through 6th rows of the surveys dataset.

##how to combine nrow() with anything because nrow will only produce a number...not contents 


```{r}
head(surveys)

surveys[-(7:34786),]
```


## Factors
```{r}
sex <- factor(c("male","female","female","male"))
levels(sex)
nlevels(sex)
sex
sex <- factor(sex, levels = c("male", "female"))
sex
```

```{r}
as.character(sex)
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(as.character(year_fct))

```

```{r}
plot(as.factor(surveys$sex))
sex <- factor(surveys$sex)
head(sex)
levels(sex)
```


```{r}
levels(sex)[1] <- "undetermined"
levels(sex)
```


##Chalenge 3
## 1.)
```{r}
levels(sex)[2] <- "female"
levels(sex)[3] <- "male"
levels(sex)
plot(as.factor(surveys$sex))
```

## 2.)
```{r}
sex <- factor(sex, levels = c("female", "male", "undetermined"))
levels(sex)
plot(sex)
```

##Using stringAsFactors = False
## Compare the difference between our data read as factor vs character.
```{r}
surveys <- read.csv("data_raw/portal_data_joined.csv", stringsAsFactors = TRUE)
str(surveys)
```

```{r}
surveys <- read.csv("data_raw/portal_data_joined.csv", stringsAsFactors = FALSE)
str(surveys)
```

## Convert the column "plot_type" into a factor
```{r}
surveys$plot_type <- factor(surveys$plot_type)
```

```{r}
animal_data <- data.frame(animal = c("dog", "cat", "sea cucumber", "sea urchin"), feel = c("furry", "squishy", "spiny"), weight = c(45, 8 , 1.1, 0.8))
```




##Challenge 4: Find the mistakes in this hand crafted data
```{r}
animal_data <- data.frame(
          animal = c("dog", "cat", "sea cucumber", "sea urchin"),
          feel = c("furry", "furry", "squishy", "spiny"),
          weight = c(45, 8, 1.1, 0.8)
          )
```



##Can you predict the class for each of the columns in the following example? Check your guesses using str(country_climate):
##    Character, Character, Number, Logistical, Logistical

##Are they what you expected? Why? Why not?
##     They werent as expected because of the errors in the data. 

##What would you need to change to ensure that each column had the accurate data type?
```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, "15"),
       northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       has_kangaroo = c(FALSE, FALSE, FALSE, 1)
       )
```

## After a few changes, the columns had the accurate data type

```{r}
country_climate <- data.frame(
       country = c("Canada", "Panama", "South Africa", "Australia"),
       climate = c("cold", "hot", "temperate", "hot/temperate"),
       temperature = c(10, 30, 18, 15),
       northern_hemisphere = c(TRUE, TRUE, FALSE, FALSE),
       has_kangaroo = c(FALSE, FALSE, FALSE, TRUE)
       )

str(country_climate)


```