---
title: "lab3b_lauren.Rmd"
output: html_document
---
## Manipulating Data, analyzing and exporting data with tidyverse
## With dplyr

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
surveys <- read_csv("data_raw/portal_data_joined.csv")
```
##Inspecting the data

```{r}
str(surveys)
```
##preview the data
```{r}
view(surveys)
```

##Selecting columns and filtering rows
```{r}
select(surveys, plot_id, species_id, weight)
```
```{r}
select(surveys, -record_id, -species_id)
```

```{r}
filter(surveys, year == 1995)
```

##Pipes: select and filter at the same time (Control, Shift, M)

```{r}
surveys2 <- filter(surveys, weight < 5)
surveys_sml <- select(surveys2, species_id, sex, weight)
```

```{r}
surveys %>% 
filter(weight < 5) %>%
select(species_id, sex, weight)
```

```{r}
surveys_sml <- surveys %>%
  filter(weight < 5) %>%
  select(species_id, sex, weight)
```


##Challenge 1: Using pipes, subset the surveys data to include animals collected before 1995 and retain only the columns year, sex, and weight.
```{r}
surveys_a95 <- surveys %>% 
  filter(year < 1995) %>% 
  select(year, sex, weight)
```


---------------------------------------------------------------------------------------------
## Mutate: to do unit conversions, or to find the ratio of values in two columns


##Challenge 2: Create a new data frame from the surveys data that meets the following criteria: contains only the species_id column and a new column called hindfoot_cm containing the hindfoot_length values converted to centimeters. In this hindfoot_cm column, there are no NAs and all values are less than 3. Hint: think about how the commands should be ordered to produce this data frame!

```{r}
surveys_chlng2 <- surveys %>%
  mutate(hindfoot_cm = hindfoot_length / 10) %>% 
  filter(!is.na(hindfoot_cm)) %>%
  filter(hindfoot_cm < 3) %>% 
  select(species_id, hindfoot_cm)
```





---------------------------------------------------------------------------------------------
##Split-apply-combine data analysis and the summarize() function: split the data into groups, apply some analysis to each group, and then combine the results

```{r}
surveys %>%
  group_by(sex) %>%
  summarize(mean_weight = mean(weight, na.rm = TRUE))
```
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight), min_weight= min(weight)) %>%
  arrange((min_weight))
```
```{r}
surveys %>%
  filter(!is.na(weight)) %>%
  group_by(sex, species_id) %>%
  summarize(mean_weight = mean(weight),
            min_weight = min(weight)) %>%
  arrange(desc(mean_weight))
```

##Counting

```{r}
surveys %>%
    count(sex, sort = TRUE)

surveys %>%
  count(sex, species)

surveys %>%
  count(sex, species) %>%
  arrange(species, desc(n))
```




##Challenge 3:

## 1.) How many animals were caught in each plot_type surveyed?

```{r}
surveys %>%
  count(plot_type)
```

## 2.) Use group_by() and summarize() to find the mean, min, and max hindfoot length for each species (using species_id). Also add the number of observations (hint: see ?n).
```{r}
surveys %>%
    filter(!is.na(hindfoot_length)) %>%
    group_by(species_id) %>%
    summarize(mean_length = mean(hindfoot_length), min_length = min(hindfoot_length), max_length = max(hindfoot_length))
```
```{r}
surveys %>%
    count(species_id)
```


## 3.) What was the heaviest animal measured in each year? Return the columns year, genus, species_id, and weight.

```{r}
surveys_maxweight <- surveys %>% 
  group_by(year, genus, species_id) %>% 
  summarize(weight = max(weight))
```
------------------------------------------------------------------------------------------

##Reshaping with Gather and Spread

```{r}
surveys_gw <- surveys %>%
  filter(!is.na(weight)) %>%
  group_by(plot_id, genus) %>%
  summarize(mean_weight = mean(weight))
```

```{r}
str(surveys_gw)
```
```{r}
surveys_spread <- surveys_gw %>%
  spread(key = genus, value = mean_weight)

str(surveys_spread)
```
```{r}
surveys_gw %>%
  spread(genus, mean_weight, fill = 0) %>%
  head()

```


##Challenge 4

##Spread the surveys data frame with year as columns, plot_id as rows, and the number of genera per plot as the values. You will need to summarize before reshaping, and use the function n_distinct() to get the number of unique genera within a particular chunk of data. It's a powerful function! See ?n_distinct for more.

```{r}
surveys_spread_chlng4 <- surveys %>%
  group_by(plot_id, year) %>%
  summarize(distinct_genera = n_distinct(genus)) %>% 
  spread(key = year, value = distinct_genera)
  head(surveys_spread_chlng4)
```


##2.) Now take that data frame and gather() it again, so each row is a unique plot_id by year combination. !!

```{r}
surveys_gather_chlng4 <- surveys_spread_chlng4 %>%
  gather(key = "year", value = "distinct_genera", -plot_id)
  head(surveys_gather_chlng4)
```
 
##3.) The surveys data set has two measurement columns: hindfoot_length and weight. This makes it difficult to do things like look at the relationship between mean values of each measurement per year in different plot types. Let's walk through a common solution for this type of problem. First, use gather() to create a dataset where we have a key column called measurement and a value column that takes on the value of either hindfoot_length or weight. Hint: You'll need to specify which columns are being gathered.

```{r}
surveys_gather2_chlng4 <- surveys %>% 
  filter(!is.na(hindfoot_length)) %>% 
  gather("measurement", "value", hindfoot_length, weight)
head(surveys_gather2_chlng4)
```



##4.) With this new data set, calculate the average of each measurement in each year for each different plot_type. Then spread() them into a data set with a column for hindfoot_length and weight. Hint: You only need to specify the key and value columns for spread(). !!


## Data Clean Up

```{r}
surveys_complete <- surveys %>%
  filter(!is.na(weight),           # remove missing weight
         !is.na(hindfoot_length),  # remove missing hindfoot_length
         !is.na(sex))                # remove missing sex
```

## Extract the most common species_id
```{r}
species_counts <- surveys_complete %>%
    count(species_id) %>% 
    filter(n >= 50)
```

## Only keep the most common species

```{r}
surveys_complete <- surveys_complete %>%
  filter(species_id %in% species_counts$species_id)
```

```{r}
dim(surveys_complete)
```

```{r}
write_csv(surveys_complete, path = "Data/surveys_complete.csv")
```


