---
title: "Lab4_lauren.Rmd"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
```


```{r}
report_03_11_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-11-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State")
```

```{r}
str(report_03_11_2020)
```

```{r}
View(report_03_11_2020)
```

```{r}
library(DT)
datatable(report_03_11_2020)
```

##Exercises: Part 1

## 3/12/2020

```{r}
report_03_12_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-12-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State")
```

```{r}
str(report_03_12_2020)
```

```{r}
view(report_03_12_2020)
```


## 9/12/2020

```{r}
report_09_12_2020 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-12-2020.csv"))
```

```{r}
str(report_09_12_2020)
```

```{r}
view(report_09_12_2020)
```


## What are the differences between 3/12/2020, the Friday before spring break 2020 and yesterday 6 months later on 9/12/2020?

##    The data from six months later now has columns for FIPS, Admin 2, Active, Combined_Key, Incidence_Rate, Case-Fatality_Ratio

## Many countries have multiple entries representing different regions (e.g. counties and states). To get the counts for a country we need to summarise the data. Use group_by() and summarize() to make a data frame (tibble) with total counts for each country.

```{r}
report_03_12_2020 %>% 
  group_by(Country_Region) %>% 
  summarize(Deaths_sum = sum(Deaths))

```

## To get counts for confirmed cases per U.S state use filter() to select US and group_by and summarize as above.

```{r}
temp <- report_03_12_2020 %>% 
  filter(Country_Region == "US") %>% 
  group_by(Province_State) %>% 
  summarize(Confirmed_sum = sum(Confirmed))
  head(temp)
```



## It will be difficult to make simple graphs with the data from all countries in the world or even the US states. Use arrange_by() and slice() to get just the top 10 countries for Deaths.

```{r}
temp <- report_03_12_2020 %>% 
  group_by(Country_Region) %>% 
  summarize(Deaths_sum = sum(Deaths)) %>% 
  arrange(desc(Deaths_sum)) %>% 
  slice(1:10)
  head(temp)
```

## R Graphics

```{r}
install.packages("tidyverse")
```


## Exercises: Part 2

## Make plots using geom_point() of the 10 countries with the confirmed cases and deaths (two separate graphs) on 3/12/2020 and 9/12/2020.

## Plot 3/12/2020
```{r}
surveys_top10_03_12 <- report_03_12_2020 %>% 
  group_by(Country_Region) %>% 
  summarize(Deaths_sum = sum(Deaths)) %>% 
  arrange(desc(Deaths_sum)) %>% 
  slice(1:10)
```

```{r}
ggplot(data = surveys_top10_03_12, mapping = aes(x = Deaths_sum, y = Country_Region)) +
    geom_point(alpha = 3)
```

## Plot 9/12/2020

```{r}
surveys_top10_09_12 <- report_09_12_2020 %>% 
  group_by(Country_Region) %>% 
  summarize(Deaths_sum = sum(Deaths)) %>% 
  arrange(desc(Deaths_sum)) %>% 
  slice(1:10)
```

```{r}
ggplot(data = surveys_top10_09_12, mapping = aes(x = Deaths_sum, y = Country_Region)) +
    geom_point(alpha = 3)
```


---------------------------------------------------------------------------------------------


## Make using geom_bar of the 10 states with the most confirmed cases and deaths (two separate graphs) on 3/12/2020 and 9/12/2020.

## 3/12/2020
```{r}
 ggplot(data = surveys_top10_03_12, mapping = aes(x = Deaths_sum,  y = Country_Region)) + 
    geom_bar(stat = 'identity') 
```

## 9/12/2020

```{r}
 ggplot(data = surveys_top10_09_12, mapping = aes(x = Deaths_sum,  y = Country_Region)) + 
    geom_bar(stat = 'identity') 
```

