---
title: "lab5_lauren.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data wrangling and graphing COVID-19 reporting data (Part 2)

```{r}
library(tidyverse)
```

```{r}
Confirmed_State_3_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-13-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State") %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```
```{r}
str(Confirmed_State_3_13)
```

```{r}
Confirmed_State_9_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv")) %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 
```
```{r}
str(Confirmed_State_9_13)
```

```{r}
setdiff(Confirmed_State_9_13$Province_State, Confirmed_State_3_13$Province_State)
```

```{r}
Confirmed_State_9_13 <- Confirmed_State_9_13 %>% 
  filter(Province_State != "Recovered") 
```

```{r}
Confirmed_State_3_13_9_13_joined <- full_join(Confirmed_State_3_13,
      Confirmed_State_9_13, by = c("Province_State"))
head(Confirmed_State_3_13_9_13_joined)
```
```{r}
tail(Confirmed_State_3_13_9_13_joined, 5)
```
```{r}
Confirmed_State_3_13_9_13_joined <- full_join(Confirmed_State_3_13,
      Confirmed_State_9_13, by = c("Province_State")) %>% 
      rename(Confirmed_3_13_2020 = "Confirmed.x", Confirmed_9_13_2020 = "Confirmed.y") %>% 
      select(-Country_Region.x, -Country_Region.y) %>% 
      replace_na(list(Confirmed_3_13_2020 = 0))
head(Confirmed_State_3_13_9_13_joined)
```
```{r}
which(is.na(Confirmed_State_3_13_9_13_joined))
```

## Switching between wide and long table formats

```{r}
Confirmed_State_3_13_9_13_joined_long <- Confirmed_State_3_13_9_13_joined %>% 
              pivot_longer(-c(Province_State),
                            names_to = "Date", values_to = "Confirmed")
```

```{r}
ggplot(Confirmed_State_3_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State))  + 
    geom_point(aes(color = Date)) 
```
## Working with the time series data
## Data wrangling
## start by loading the most recent time series data for confirmed cases

```{r}
time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")
```


```{r}
head(time_series_confirmed)
```
```{r}
time_series_confirmed_long <- time_series_confirmed %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 
```

```{r}
head(time_series_confirmed_long)
```
```{r}
 download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", 
               destfile = "Data/time_series_covid19_deaths_global.csv")
```

```{r}
time_series_deaths <- read_csv("Data/time_series_covid19_deaths_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")
```

```{r}
time_series_deaths_long <- time_series_deaths %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Deaths") 
head(time_series_deaths_long)
```
## Joining confirmed and deaths data: by mergering the names for the Province_State, Country_Region and Date columns using unite

```{r}
time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
head(time_series_confirmed_long)
```
## eliminate redundancy: so that we just have the Key and Deaths columns
 
```{r}
time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)
```


```{r}
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key")) %>% 
              select(-Key)
head(time_series_long_joined)
```
```{r}
which(is.na(time_series_long_joined$Confirmed))
```

```{r}
which(is.na(time_series_long_joined$Deaths))
```

```{r}
library(lubridate)
```

```{r}
time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
```

## Confirmed and Deaths are count data and there may be times when we want to plot both on the same graph

```{r}
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date),
               names_to = "Report_Type", values_to = "Counts")
head(time_series_long_joined_counts)
```
## Making Graphs from time series data

```{r}
time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Deaths)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Deaths")
```
```{r}
time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region %in% c("China","Japan", "Korea, South",
                                "Italy","Spain", "US")) %>% 
    ggplot(aes(x = Date,  y = Deaths)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths") +
    facet_wrap(~Country_Region, ncol=2, scales="free_y")
```

```{r}
time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("China","France","Italy", 
                                "Korea, South", "US")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths")
```

```{r}
time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = log2(Counts), fill = Report_Type, color = Report_Type)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Cases")
```

## Exercises

## 1.) Finish the Data Visualization tutorial and add it your GitPage web site
##    See lab5_datavisual_lauren.Rmd

## 2.) Revise the above example for joining tables using 6/13/2020 and 9/13/2020. Plot the data as a bar plot. Make sure your graph looks nice in the report by adjusting the height in the R code chunk header.


```{r}
time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Counts, fill = Report_Type, color = Report_Type)) + 
    geom_bar(stat='identity') +
    ggtitle("US COVID-19 Cases")
```

## 3.) Add more informative x and y labels and a title to the above graph in Ex2.


```{r}
time_series_long_joined_counts %>% 
  group_by(Country_Region, Report_Type, Date) %>% 
  summarise(Counts = sum(Counts)) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Counts, fill = Report_Type, color = Report_Type)) + 
    geom_bar(stat='identity') +
        labs(title = "US COVID-19 Cases",
        x = "Date",
        y = "Count (logarithmic)") +
    theme_bw() +
    theme(axis.text.x = element_text(colour = "grey20", size = 12,
                                     hjust = 0.5, vjust = 0.5),
          axis.text.y = element_text(colour = "grey20", size = 12),
          text = element_text(size = 16))
```

## 4.) Using the time series data. Make a plot of the total number of confirmed deaths per day worldwide

```{r}
time_series_long_joined %>% 
  group_by(Date) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  ggplot(aes(x = Date,  y = Deaths)) +
  geom_point() +
  geom_line() +
  ggtitle("Wordwide COVID-19 Deaths")
```



## 5.) Use Mutate to make a new column with deaths/confirmed cases (Hint: this might be best done after joining the tables, but before pivoting longer).

```{r}
time_series_ratio <- time_series_long_joined %>% 
  mutate(ratio = Deaths/Confirmed)
```

## 6.) Plot US deaths/confirmed cases per day.

```{r}
time_series_ratio <- time_series_long_joined %>% 
  filter (Country_Region == "US") %>%
  mutate(ratio = Deaths / Confirmed) %>% 
  group_by(Date) %>% 
  summarize(ratio_sum = sum(ratio)) %>% 
  ggplot(aes(x = Date,  y = ratio_sum)) +
  geom_point() +
  geom_line() +
  ggtitle("US COVID-19 Deaths")
time_series_ratio
```



## 7.) Make a single graph with from the times series data with the 10 countries with the hightest death total.


```{r}
time_series_top10 <- time_series_long_joined %>%
  filter(Date == "2020-11-18") %>% 
  group_by(Country_Region) %>% 
  summarize(Deaths_sum = sum(Deaths)) %>% 
  arrange(desc(Deaths_sum)) %>% 
  slice(1:10)
top10 <- time_series_top10$Country_Region
```

```{r}
time_series_long_joined %>% 
  filter(Country_Region %in% top10) %>%
  group_by(Country_Region, Date) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  ggplot(aes(x = Date, y = Deaths, fill = Country_Region, color = Country_Region)) +
  geom_point() +
  geom_line() +
  ggtitle("Top 10 Countries with COVID-19 Deaths")
```


## 8.) Using Facet_Wraps make a graph from the times series data with the 10 countries with the hightest death total. Format the table to look nice in the report by changing the number of rows in your graph and/or adjusting the height in the R code chunk header.

```{r}
time_series_long_joined %>% 
  filter(Country_Region %in% top10) %>%
  group_by(Country_Region, Date) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  ggplot(aes(x = Date, y = Deaths, fill = Country_Region, color = Country_Region)) +
  geom_point() +
  geom_line() +
  ggtitle("Top 10 Countries with COVID-19 Deaths") +
  facet_wrap(facets = vars(Country_Region)) +
  labs(title = "Top 10 Countries with COVID-19 Deaths", x = "Date of Observation", y = "Death Count") +
  theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 12),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16))
```


## 9.) Using Facet_Wraps make a graph from the times series data for confirmed cases for all US states and territories. Format the table to look nice in the report by changing the number of rows in your graph and/or adjusting the height in the R code chunk header.

```{r}
time_series_confirmed_long %>% 
  filter(Country_Region == "US") %>%
  group_by(Province_State, Date) %>% 
  summarize(Confirmed = sum(Confirmed))
```  
  
  ggplot(aes(x = Date, y = Confirmed)) +
  geom_point() +
  geom_line() +
  facet_wrap(facets = vars(Country_Region)) +
  labs(title = "U.S. Confirmed COVID-19 Cases", x = "Date of Observation", y = "Confirmed Count (Logarithmic)") +
  theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 9),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16))
```

## 10.) Apply a theme other than the one used in the Data Carpentries tutorial to the graph in Ex8



```{r}
time_series_new_theme <- time_series_long_joined %>% 
  filter(Country_Region %in% top10) %>%
  group_by(Country_Region, Date) %>% 
  summarize(Deaths = sum(Deaths)) %>% 
  ggplot(aes(x = Date, y = Deaths, fill = Country_Region, color = Country_Region)) +
  geom_point() +
  geom_line() +
  ggtitle("Top 10 Countries with COVID-19 Deaths") +
  facet_wrap(facets = vars(Country_Region)) +
  labs(title = "Top 10 Countries with COVID-19 Deaths", x = "Date of Observation", y = "Death Count") +
  theme_dark() +
  theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 12),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16))
time_series_new_theme
```

